Bias correction in GLMs

### 2021-07-20 bias_correction.rmd
Population averaging seems to work in a simple case

```{r sim_simple}
N <- 1e5
beta0 <- 0.7
betaA <- 0.2
betaW <- 0.5

age_max <- 1
age_min <- 0.2
# age <- runif(N, age_min, age_max)
age <- rnorm(N, age_max, age_max)

@@

## Pop. average
simple_vareff_age_pop <- popavefun(simple_mod, "age", "wealthindex", modelname = "Pop. ave") 
binned_df <- simple_vareff_age_pop$check

vareff_age <- simple_vareff_age
vareff_age$preds <- do.call("rbind", list(vareff_age$preds, simple_vareff_age_pop$preds))
age_plot <- (plot(vareff_age)
 	+ labs(y="Prob. of improved \n service", colour="Model")
	+ geom_hline(yintercept=true_prop, lty=2, colour="grey")
	+ geom_point(data=binned_df, aes(x=age, y=status, color="binned"))
	+ scale_colour_manual(values=c("black", "blue", "red"))
	+ theme(legend.position="bottom")
)
```

@@

```{r simple_vareff_wealthindex, fig.cap = "A comparison of population averaged and varpred-based predictions. We also apply bias-adjustemnt to the `vapred` predictions. For `age`, we implement the naive approach to compute the predictions in `popavefun` function and then implement the same in `vareffects` so as to use the centering machineries. In both cases, the population averaging and varpred gives similar estimates. The estimated population average is very close to the observed in the case of `wealthindex` but slightly higher in the case of `age` for the unadjusted estimates (but very low for bias-adjusted). , see previous paragraph."}
# Wealth index
## varpred 
simple_vareff_wealthindex <- varpred(simple_mod, "wealthindex", isolate=FALSE, modelname="varpred")

## Pop. average
simple_vareff_wealthindex_pop <- popavefun(simple_mod, "wealthindex", "age", modelname="Pop. ave")
binned_df <- simple_vareff_wealthindex_pop$check

vareff_wealthindex <- simple_vareff_wealthindex
vareff_wealthindex$preds <- do.call("rbind", list(vareff_wealthindex$preds, simple_vareff_wealthindex_pop$preds))
wealthindex_plot <- (plot(vareff_wealthindex)
 	+ labs(y="", colour="Model")
	+ geom_hline(yintercept=true_prop, lty=2, colour="grey")
	+ geom_point(data=binned_df, aes(x=wealthindex, y=status, color="binned"))
	+ scale_colour_manual(values=c("black", "blue", "red"))
	+ theme(legend.position="bottom")
)

@@

\section{Random effect model}

```{r sim_randeff}
# Simulation parameters
nHH <- 1000		# Number of HH (primary units) per year

nyrs <- 10	# Number of years to simulate
yrs <- 2000 + c(1:nyrs) # Years to simulate
N <- nyrs * nHH

## HH random effect sd
hhSD <- 0.5

# Generate dataset template
temp_df <- (data.frame(hhid = rep(c(1:nHH), each = nyrs)
		, years = rep(yrs, nHH)
		, age = runif(n=N, age_min, age_max)
		, wealthindex = rnorm(n = N, 0, 1)
	)
)

# Simulate HH-level random effects (residual error)
hhRE <- rnorm(nHH, hhSD)
temp_df$hhRE <- hhRE[temp_df$hhid]

sim_df <- (temp_df
	%>% mutate(eta = beta0 + betaA * age + betaW * wealthindex + hhRE
		, status = rbinom(N, 1, plogis(eta))
	)
	%>% select(-eta)
)
true_prop_reff <- mean(sim_df$status)
print(true_prop)
print(head(sim_df, 50))
```

\subsection{Fit model}

```{r model_reff}
reff_mod <- glmmTMB(status ~ age + wealthindex + (1|hhid)
	, data = sim_df
	, family = binomial(link = "logit")
)
```

```{r reff_coef}
## Tidy coef estimates
reff_coef_df <- (broom.mixed::tidy(reff_mod, conf.int=TRUE)
	%>% mutate(term = gsub("\\(|\\)", "", term))
	%>% filter(effect=="fixed")
)
print(reff_coef_df)

reff_coef_plot <- (plotEsize(reff_coef_df) 
	+ geom_point(data=true_beta_df, aes(x=term, y=estimate, colour="Truth"))
	+ labs(colour="Type")
	+ scale_colour_manual(values=c("black", "blue"))
)
print(reff_coef_plot)
```

Variable effect plots

```{r reff_vareffects_age} 
## varpred way
reff_vareff_age <- varpred(reff_mod, "age", isolate=FALSE, modelname="varpred")

## Pop. average
reff_vareff_age_pop <- varpred(reff_mod, "age", isolate=TRUE, pop.ave=TRUE, modelname = "Pop. ave")

## Bias adjust
reff_vareff_age_adjust <- varpred(reff_mod, "age", isolate=TRUE, bias.adjust=TRUE, modelname = "Bias adjust")

vareff_age <- reff_vareff_age
vareff_age$preds <- do.call("rbind", list(vareff_age$preds, reff_vareff_age_pop$preds, reff_vareff_age_adjust$preds))
age_plot <- (plot(vareff_age)
 	+ labs(y="Prob. of improved \n service", colour="Model")
	+ geom_hline(yintercept=true_prop_reff, lty=2, colour="grey")
	+ scale_colour_manual(values=c("black", "blue", "red"))
	+ theme(legend.position="bottom")
)

```{r reff_vareff_wealthindex, fig.cap = "A comparison of population averaged, varpred-based and bias-adjusted predictions. For each of the predictors, the population averaging and varpred gives similar estimates, and slightly off the truth. However, when we apply bias-adjustment, the estimates are very precise (i.e., close to the truth)."}
# Wealth index
## varpred 
reff_vareff_wealthindex <- varpred(reff_mod, "wealthindex", isolate=TRUE, modelname="varpred")

## Pop. average
reff_vareff_wealthindex_pop <- varpred(reff_mod, "wealthindex", isolate=TRUE, pop.ave=TRUE, modelname="Pop. ave")

## Bias adjust
reff_vareff_wealthindex_adjust <- varpred(reff_mod, "wealthindex", isolate=TRUE, bias.adjust=TRUE, modelname="Bias adjust")

vareff_wealthindex <- reff_vareff_wealthindex
vareff_wealthindex$preds <- do.call("rbind", list(vareff_wealthindex$preds, reff_vareff_wealthindex_pop$preds, reff_vareff_wealthindex_adjust$preds))
wealthindex_plot <- (plot(vareff_wealthindex)
 	+ labs(y="", colour="Model")
	+ geom_hline(yintercept=true_prop_reff, lty=2, colour="grey")
	+ scale_colour_manual(values=c("black", "blue", "red"))
	+ theme(legend.position="bottom")
)

>>>>

```{r sim_simple}
N <- 1e4
beta0 <- 0.7
betaA <- 0.2
betaW <- 0.5

age_max <- 1
age_min <- 0.2
age <- runif(N, age_min, age_max)
# age <- rnorm(N, age_max, age_max)

@@

## Pop. average
simple_vareff_age_pop <- varpred(simple_mod, "age", isolate=FALSE, pop.ave=TRUE, modelname="Pop. ave")
# simple_vareff_age_pop <- popavefun(simple_mod, "age", "wealthindex", modelname = "Pop. ave") 
# binned_df <- simple_vareff_age_pop$check

vareff_age <- simple_vareff_age
vareff_age$preds <- do.call("rbind", list(vareff_age$preds, simple_vareff_age_pop$preds))
age_plot <- (plot(vareff_age)
 	+ labs(y="Prob. of improved \n service", colour="Model")
	+ geom_hline(yintercept=true_prop, lty=2, colour="grey")
	+ scale_colour_manual(values=c("black", "blue", "red", "green"))
	+ theme(legend.position="bottom")
)
```

@@

```{r simple_vareff_wealthindex, fig.cap = "A comparison of population averaged and varpred-based predictions. In the case of biased predictions (age), population averaging gives better estimates. On the other hand, for wealthindex, the estimates varpred and population estimates are somehow similar."} 
# Wealth index
## varpred 
simple_vareff_wealthindex <- varpred(simple_mod, "wealthindex", isolate=FALSE, modelname="varpred")

## Pop. average
simple_vareff_wealthindex_pop <- varpred(simple_mod, "wealthindex", isolate=FALSE, pop.ave=TRUE, modelname="Pop. ave")
# binned_df <- simple_vareff_wealthindex_pop$check

vareff_wealthindex <- simple_vareff_wealthindex
vareff_wealthindex$preds <- do.call("rbind", list(vareff_wealthindex$preds, simple_vareff_wealthindex_pop$preds))
wealthindex_plot <- (plot(vareff_wealthindex)
 	+ labs(y="", colour="Model")
	+ geom_hline(yintercept=true_prop, lty=2, colour="grey")
#	+ geom_point(data=binned_df, aes(x=wealthindex, y=status, color="binned"))
	+ scale_colour_manual(values=c("black", "blue", "red"))
	+ theme(legend.position="bottom")
)

@@
